---
title: "Code"
author: "Eric Chen, Junhan Li, & Daniel Fredin"
output: pdf_document
---

```{r include=FALSE, echo=TRUE}
# Libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(splitstackshape)
library(readr)

library(car)
library(lmtest)
library(rcompanion)
library(olsrr)

library(maps)
library(usmap)
```

```{r include=FALSE, echo=TRUE}
# Load datasets
attorneys <- read.csv("attorneys.csv")
attorneys_time <- read.csv("attorneytimeentries.csv")
categories <- read.csv("categories.csv")
clients <- read.csv("clients.csv")
questionposts <- read.csv("questionposts.csv")
questions <- read.csv("questions.csv")
statesites <- read.csv("statesites.csv")
subcategories <- read.csv("subcategories.csv")
```

#### Data Cleaning

```{r}
# Remove unwanted columns for all datasets
attorneys <- attorneys %>%
  select(-Id, -CreatedUtc, -City, -PostalCode)

attorneys_time <- attorneys_time %>%
  select(-Id, -EnteredOnUtc, -TimeEntryUno)

categories <- categories %>%
  select(-Id)

clients <- clients %>%
  select(-Id, -PostalCode, -CreatedUtc, -InvestmentsBalance)

questionposts <- questionposts %>%
  select(-Id, -CreatedUtc)

questions <- questions %>%
  select(-Id, -AskedOnUtc, -TakenOnUtc, -ClosedOnUtc, -LegalDeadline, -ClosedByAttorneyUno)

statesites <- statesites %>%
  select(-Id)

subcategories <- subcategories %>%
  select(-Id)
```

```{r}
# Convert characters to numeric in client dataset
clients$Age <- as.integer(clients$Age)
clients$NumberInHousehold <- as.integer(clients$NumberInHousehold)
clients$AnnualIncome <- as.numeric(clients$AnnualIncome)
clients$AllowedIncome <- as.numeric(clients$AllowedIncome)
clients$CheckingBalance <- as.numeric(clients$CheckingBalance)
clients$SavingsBalance <- as.numeric(clients$SavingsBalance)
```

```{r}
# rename columns to match other datasets
questions <- questions %>%
  rename("ClientUno" = "AskedByClientUno", "AttorneyUno" = "TakenByAttorneyUno")
```

```{r}
# Merge into two dataframes

# df1 is all datasets except attorney and attorney_time
df1 <- left_join(clients, questions, join_by("StateAbbr", "ClientUno")) %>% 
  left_join(., categories, join_by("StateAbbr", "CategoryUno", "Category")) %>%
  left_join(.,subcategories, join_by("StateAbbr", "CategoryUno", "Subcategory", "SubcategoryUno")) %>%
  left_join(., questionposts, join_by("StateAbbr", "QuestionUno")) %>%
  left_join(., statesites, join_by("StateAbbr", "StateName"))

# df2 is the two attorney datasets
df2 <- left_join(attorneys, attorneys_time, join_by("StateAbbr", "AttorneyUno")) 
```

```{r}
# Select the independent variables we want to observe
# Remove all NA

data_overall <- df1 %>%
  select(StateAbbr, Age, NumberInHousehold, MaritalStatus, AnnualIncome, SavingsBalance, CheckingBalance, AllowedIncome, Category, Subcategory, ClientUno) %>%
  distinct(ClientUno, .keep_all = TRUE) %>%
  select(-ClientUno) %>%
  na.omit(df1)
```

# Change the subcategory to binary categorical data 
Divorce mentioned entry is represented by 1, otherwise 0. 
```{r}
data_overall$Subcategory <- ifelse(str_detect(data_overall$Subcategory, "Divorce"), 1, 0)
```


Datasets: 
  - Statesites
    - StateAbbr
    - AllowedAssets
    - BaseIncomeLimit
    - PerHouseholdMemberIncomeLimit
    - IncomeMultiplier 
  - Client
    - StateAbbr
    - Prettymuch the entire dataset 
  - Categories
    - Category: Family and Children 
   -nSubcategories: everything related to divorces
   
   
   
   
# Visualization 1: Investigating the Top 15 Subcategories of Asked Questions
```{r}
top_subcats <- questions %>%
  group_by(Subcategory) %>%
  summarise(num_subcats = n()) %>%
  ungroup()

top_subcats <- top_subcats %>%
  arrange(desc(num_subcats)) %>%
  head(15)

ggplot(top_subcats, aes(x = num_subcats, y = reorder(Subcategory, num_subcats))) +
  geom_bar(stat = "Identity", fill = "#00A19B") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 15 Subcategories",
       x = "Number of Clients",
       y = "Subcategory")

```
Interpretation:
  For our project, we want to started off with investigating the most frequently 
asked legal questions subcategories so we can define a most common question type
based on the data set we have at hand and use it as the theme of our research 
question. 

  Based on our side way bar chart, it is clear to see that out of the top 15 
subcategories, among the top three categories of questions asked by clients on 
the online platform, two of them are related to divorce.The question of subcategory 
Family/Divorce/Custody has the highest frequency among the top 15, and it has been
asked by almost twice as many clients than the second subcategory of others, which
is showing that the divorce related questions should be treated with more 
preparation when it comes to training volunteers.  

  This is intriguing to us as it begs for answers to the questions such as 
"What are the major determinants of divorce?","Does the clients' backgrounds
affect their tendency to ask divorce related question on the ABA online platform?"
,and most importantly, "How do we prepare our volunteers to address these divorce
related questions?"
  

# Visualization 2: US Map of Divorce Related Questions Distribution
```{r setup, include=TRUE}
# Create a data set with divorce related questions frequency per state
contingency <- table(data_overall$StateAbbr,data_overall$Subcategory)
divorce_distrib <- as.data.frame(contingency)
divorce_distrib[divorce_distrib$Var2 != 0, ]  

# Change the column name var1 to state, so we can use plot_usmap
colnames(divorce_distrib)[which(names(divorce_distrib) == "Var1")] <- "state"

# state_divorce <- 
plot_usmap(data = divorce_distrib, values = "Freq", color = "red") + 
  scale_fill_continuous(
    low = "white", high = "red", name = "Frequency of Divorce Related Questions"
    , label = scales::comma
  ) + theme(legend.position = "right")

```
Certain States such as: are shown in gray because out of confidentiality and per
state law, they are not allowed to provided clients; information. 

However, the US map plot of distribution of divorce rate 


Rubbish Code:(Keep if for now)
ggplot(data=divorce_distrib, aes(x=long, y=lat, fill=region, group=group)) + 
  geom_polygon(color = "white", fill="gray") + 
  guides(fill=FALSE) + 
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
  ggtitle('Distribution of the Marriage Issues Across United States') + 
  coord_fixed(1.3)
  
# Visualization 3: Financial Status Correlation with Divorce Rate
We need to use that facet, wrap, colored shit here. 


#### Testing models

```{r}
data_overall$Subcategory <- ifelse(str_detect(data_overall$Subcategory, "Divorce"), 1, 0)
```

# Model 1
```{r}

model <- glm(factor(Subcategory) ~ . -Category -StateAbbr, data = data_overall, family = 'binomial')
summary(model)
```
Assumes that there is no severe multicollinearity
Assume 


